Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-18 paper.tex
chore: more on evaluation

The programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}, 
 a core language that allows the expression of collective behaviours through the composition of \emph{functions} operating on fields.
%
Field calculus includes main operators for expressing \emph{spatio-temporal} compotations that allow for 
\emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 
\emph{ii)} the exchange of data with neighbouring fields, where received data is accessed by adjacent fields; and 
\emph{iii)} the conditional division of computation into distinct domains of collective computation. 
%
%We avoid presenting here the full details of these operators since they are not essential to the contribution. 
For further elaboration on the actual field calculus, please refer to the work of~\cite{DBLP:journals/jlap/ViroliBDACP19}.

@@

\lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
\subsection{Graph Neural Networks}
%Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
%  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
%  \lukas{more examples? }
%%
%Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio), 
 \ac{GNN} is a novel neural network model used to process graph-structured data with deep learning approaches.
%
Let $G = (V, E)$ a graph  
where $E$ defines the neighbourhood relations for each participating node, 
and $V$ identifies the nodes present in the graph. %at that time.  \lukas{no time in the definition}
% 
Each node $v \in V$ is associated with an observation (or feature set) $f_v$.
For the sake of simplicity, 
 we thereafter describe $G_f$ as a graph that contains the feature set $f_v$ for each node $v \in V$.
 Note that, when we refer to $G_f$ and $G_o$ we are referring to the same graph $G$ but with different node features.
 Also, to access the feature set $f_v$ of a node $v \in V$ we use the notation $f_v$ or $G_f[v]$.
%\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
%
Given $f_v$, %a feature vector associated with each node $v \in V$, 
 the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$. The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
 In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
  and then combining it with the node's current embedding $h_v$.
 The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased
Formally, a \ac{GNN} can be defined by two phases:
\begin{equation}
a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}_G(v)\right\}\right)   
\end{equation}

\begin{equation}
h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
\end{equation}
where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
 $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
%
$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
%
The differential part comes into play in the $\operatorname{COMBINE}$ function, 
 which is usually a differentiable function such as a neural network layer.
%
$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
  and it could be a simple sum, max or sum of products. More complex aggregation functions are available~\cite{pellegrini2020learning}.
%
Thereafter, we express the application of a \ac{GNN} to a graph $G_f$ as:
\begin{equation}
    GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
\end{equation} 
This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}

\acp{GNN} are used in several application areas such as social network analysis, chemistry, and physics.
In this paper, we use GNNs to learn a local behaviour for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
\subsection{Many-Agent Reinforcement Learning}
%\ga{plan: we could discuss briefly about standard approach (few agents), and mean-field approach (many agents). 
%This will lead to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}

@@

 where the set of agents $N \gg 2$ and each agent is interchangeable and indistinguishable.

@@

In this section, we discuss the components involved in our proposed solution (i.e., architecture) and how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics). Finally, we will detail the learning algorithm designed and used to synthesise the policy (i.e., learning algorithm).

@@

 \ga{todo: put pseudo code to express this dynamics}

@@

\lukas{should stay here but this is the explicit implementation of the field:}
Each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
%
The program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
This can be expressed in Scafi as follows:
\begin{lstlisting}[mathescape=true]
val source = maxHood(nbr(sense($\zeta$))) == sense($\zeta$)
G(source, Point3D.Zero, _ + nbrVector())
\end{lstlisting}
where \texttt{maxHood} is a function that returns the maximum value of the neighbourhood, 
 \texttt{nbr} is a function that returns the value of the neighbourhood, 
\texttt{sense} is a function that returns the value of the sensor, and
\texttt{nbrRange} is a function that returns an approximate direction for each agent in the neighbourhood.
%
this value will then be fed into the \ac{GNN} to compute the action to be performed.
 %\lukas{not sure if this paragraph should be its own subsection. or at least it needs  different title.}

>>>>

The programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}, 
 a core language that allows the expression of collective behaviours through the composition of \emph{functions} operating on fields.
%
Field calculus includes main operators for expressing \emph{spatio-temporal} compotations that allow for 
\emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 
\emph{ii)} the exchange of data with neighbouring fields (typically referred as \lstinline|nbr|), where received data is accessed by adjacent fields; and 
\emph{iii)} the conditional division of computation into distinct domains of collective computation. 
%
%We avoid presenting here the full details of these operators since they are not essential to the contribution. 
For further elaboration on the actual field calculus, please refer to the work of~\cite{DBLP:journals/jlap/ViroliBDACP19}.

@@

\lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
\subsection{Graph Neural Networks}
%Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
%  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
%  \lukas{more examples? }
%%
%Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio), 
 \ac{GNN} is a novel neural network model used to process graph-structured data with deep learning approaches.
%
Let $G = (V, E)$ a graph  
where $E$ defines the neighbourhood relations for each participating node, 
and $V$ identifies the nodes present in the graph. %at that time.  \lukas{no time in the definition}
% 
Each node $v \in V$ is associated with an observation (or feature set) $f_v$.
For the sake of simplicity, 
 we thereafter describe $G_f$ as a graph that contains the feature set $f_v$ for each node $v \in V$.
 Note that, when we refer to $G_f$ and $G_o$ we are referring to the same graph $G$ but with different node features.
 Also, to access the feature set $f_v$ of a node $v \in V$ we use the notation $f_v$ or $G_f[v]$.
%\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
%
Given $f_v$, %a feature vector associated with each node $v \in V$, 
 the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$. 
 The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
%
In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
 and then combining it with the node's current embedding $h_v$.
%
The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. 
 With each traversal of an edge $e \in E$ from $v$, the number of layers increased
Formally, a \ac{GNN} can be defined by two phases:
\begin{equation}
a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}_G(v)\right\}\right)   
\end{equation}

\begin{equation}
h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
\end{equation}
where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
 $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
%
$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
%
The differential part comes into play in the $\operatorname{COMBINE}$ function, 
 which is usually a differentiable function such as a neural network layer.
%
$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
 and it could be a simple sum, max or sum of products. 
 More complex aggregation functions are available~\cite{pellegrini2020learning}.
%
Thereafter, we express the application of a \ac{GNN} to a graph $G_f$ as:
\begin{equation}
GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
\end{equation} 
This formulation allows \acp{GNN} to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
%\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}

\acp{GNN} are used in several application areas such as social network analysis, chemistry, and physics.
 In this paper, we use \acp{GNN} to learn a local behaviour for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
\subsection{Many-Agent Reinforcement Learning}
%\ga{plan: we could discuss briefly about standard approach (few agents), and mean-field approach (many agents). 
%This will lead to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}

@@

 where the set of agents is large ($N \gg 2$) and each agent is interchangeable and indistinguishable.

@@

In this section, we discuss the components involved in our proposed solution (i.e., architecture) and how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics). Finally, we will detail the learning algorithm designed and used to synthesise the policy.

@@

 %\ga{todo: put pseudo code to express this dynamics}

@@

\lukas{should stay here but this is the explicit implementation of the field:}
Each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
%
The program $\Gamma$ in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
This can be expressed in Scafi as follows:
\begin{lstlisting}[mathescape=true]
val source = maxHood(nbr(sense($\zeta$))) == sense($\zeta$)
G(source, Point3D.Zero, _ + nbrVector())
\end{lstlisting}
where \texttt{maxHood} is a function that returns the maximum value of the neighbourhood, 
 \texttt{nbr} is a function that returns the value of the neighbourhood, 
\texttt{sense} is a function that returns the value of the sensor, and
\texttt{nbrRange} is a function that returns an approximate direction for each agent in the neighbourhood.
%
This value will then be fed into the $\pi_{GNN}$ to compute the action to be performed.
 %\lukas{not sure if this paragraph should be its own subsection. or at least it needs  different title.}
